---
title: "Untitled"
author: "lbian01"
date: "09/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load lib

```{r lib, include=FALSE}
library(tidyverse)
library(tidyquant)
library(moments)
library(tseries)
library(bizdays)
```


###################################################################
# read the S&P 500 Emini time series data 
spEminiTS=read.csv("./sampleData/spxEminiEx20200619.csv", header=T,as.is = T,check.names=FALSE)
spEminiTS$TradeDate=as.Date(spEminiTS$TradeDate, format = "%d/%m/%Y")


```{r ReadData}
spEminiTS=read.csv("./sampleData/spxEminiEx20200619.csv", header=T,as.is = T,check.names=FALSE)

spEminiTS$TradeDate <-
  spEminiTS$TradeDate %>% 
  as.Date("%d/%m/%Y")

```



# basis 
spEminiTS=merge(x=spEminiTS,y=underlying[c('TradeDate','SPX')],by="TradeDate",all.x=TRUE)

spEminiTS$Basis=spEminiTS$`19/06/2020`-spEminiTS$SPX
spEminiTS$BasisP=spEminiTS$SPX/spEminiTS$`19/06/2020`-1

plot(spEminiTS$TradeDate,spEminiTS$Basis, type="l")
plot(spEminiTS$TradeDate,spEminiTS$BasisP, type="l")




```{r}

```
